\textls[-15]{\thetitle \mbox{(Technical Report)}}

### 2022-10-11 syntax.tex
Should mention how computational axiom etc is modelled.
cite "explicit subst" paper

We highlight some specialities of declarative style in QIIT that differ from a
conventional setting: 
(1) This declarative formulation of type theory does not
have an operational semantics but only judgemental equality (e.g. $\beta$-reduction is now $\beta$-equality).\footnote{Later we will prove
canonicity which can imply a certain level of computational ability of
our theory} 
(2) We directly use the equality from QIIT to represent judgemental equality in our syntax, relieve from us the burden to explicit mentioning equivalence closure; however, this also mean that, \textit{every piece of the syntax data}, including all kinds of judgements (QIIT type) and their derivations will have equality information upon them  
%We will recover canonicity afterwards, which implies the computational ability of our theory. 
% (3) We are dealing with quotient data, so we are actually manipulating
% \textit{equivalence classes of terms} instead of one representative (term) from the equivalence class.
(3) Instead of using meta-level substitution, we use explicit
substitution and de Bruijn indices in our intrinsic formulation. This is
also called \textit{substitution calculus} in the literature and is favoured due
to its closedness to categorical semantic and easir to handle in mechanization. We will go through these details one-by-one.

@@

In our metatheory, for simplicity we only formulate non-dependent eliminators, i.e.,
\textit{recursors}.
%\YZ{Do dependent eliminators pose technical difficulties, or do you avoid them solely because they require more verbiage?}\EDJreply{I don't think there is technical difficulties to do that. The formulation of course will be much more horrifying if use dependent eliminator here. } 
Like our plugin implementation, each recursor is constructed by a
linkage containing handlers dealing with each constructor, and code
reuse for recursor is delegated to linkage inheritance.
This greatly simplifies the metatheory development.

>>>>

We highlight some specialities of declarative style in QIIT that differ from a
conventional setting: 
(1) This declarative formulation of type theory does not
have an operational semantics but only judgemental equality (e.g. $\beta$-reduction is now $\beta$-equality).\footnote{Later we will prove
canonicity which can imply a certain level of computational ability of
our theory} 
(2) We directly use the equality from QIIT to represent judgemental equality in our syntax, relieve from us the burden to explicit mentioning equivalence closure; however, this also mean that, \textit{every piece of the syntax data}, including all kinds of judgements (QIIT type) and their derivations will have equality information upon them  
%We will recover canonicity afterwards, which implies the computational ability of our theory. 
% (3) We are dealing with quotient data, so we are actually manipulating
% \textit{equivalence classes of terms} instead of one representative (term) from the equivalence class.
(3) Instead of using meta-level substitution, we use explicit
substitution~\cite{abadi1989subst} and de Bruijn indices in our intrinsic formulation. This is
also called \textit{substitution calculus} in the literature and is favoured due
to its closedness to categorical semantic and easir to handle in mechanization. We will go through these details one-by-one.

@@

In our metatheory, for simplicity we only formulate non-dependent eliminators, i.e.,
\textit{recursors}.\YZ{Metatheory should mention how computational axioms about case handlers, partial recursors, and axioms over partial recursors are dealt with in the metatheory.}  
%\YZ{Do dependent eliminators pose technical difficulties, or do you avoid them solely because they require more verbiage?}\EDJreply{I don't think there is technical difficulties to do that. The formulation of course will be much more horrifying if use dependent eliminator here. } 
Like our plugin implementation, each recursor is constructed by a
linkage containing handlers dealing with each constructor, and code
reuse for recursor is delegated to linkage inheritance.
This greatly simplifies the metatheory development.
